---
title: "p8105_hw1_ly2633"
author: "Leila Yan"
date: "2024-09-13"
output: github_document
---

```{r}
# load the libraries
library(tidyverse)
```


## Problem 1
```{r}
# Load the dataset
data("penguins", package = "palmerpenguins")

# Calculate the mean flipper length
mean_flipper_length <- penguins %>%
  pull(flipper_length_mm) %>%
  mean(na.rm = TRUE)
```

Write a short description of the penguins dataset (not the penguins_raw dataset) using inline R code. In your discussion, please include:
- the data in this dataset, including names / values of important variables
- the size of the dataset (using nrow and ncol)
- the mean flipper length

Answer:
This dataset consists of 8 variables and they are: species, island, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g, sex, and year (ncol=8). It contains 344 entries (nrow=344). It provides us with information about different species of penguins (eg.Adelie, Gentoo, and Chinastrap) and islands, their sex, year, as well as some measurements of their weight, flipper length, bill length and bill depth. From the dataset, we can also see there are some missing data, as NA. The mean flipper length is 200.915 mm.



Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x); color points using the species variable (adding color = ... inside of aes in your ggplot code should help).
Export your first scatterplot to your project directory using ggsave.

```{r}
# Create a scatterplot
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(
    title = "Scatterplot of Flipper Length vs Bill Length",
    x = "Bill Length (mm)",
    y = "Flipper Length (mm)"
  ) +
  theme_minimal()

# Export my scatterplot to my project directory
ggsave("scatterplot_flipper_vs_bill.png")
```
## Problem 2

This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

Create a data frame comprised of:

a random sample of size 10 from a standard Normal distribution
a logical vector indicating whether elements of the sample are greater than 0
a character vector of length 10
a factor vector of length 10, with 3 different factor “levels”
Try to take the mean of each variable in your dataframe. What works and what doesn’t?
Hint: to take the mean of a variable in a dataframe, you need to pull the variable out of the dataframe. Try loading the tidyverse and using the pull function.
```{r}
# Create data frame
problem2_df = tibble (
  randomsample= rnorm(10),
  vec_char = c ("cancer", "recurrence", "demographics", "environment", "genetic", "age", "sex", "ethnicity", "comorbidity", "medication"),
  vec_logical= rnorm(10) > 0,
  vec_factor = factor(c("small", "medium", "large", "small", "small", "medium", "large", "small", "large", "medium"))
)

problem2_df

# taking the mean of each variable
# a randomsample
randomsample <- problem2_df %>%
  pull(randomsample) %>%
  mean(na.rm = TRUE)

#vec_logical
mean_logical <- problem2_df %>%
  pull(vec_logical) %>%
  mean(na.rm = TRUE)


#vec_factor
mean_factor <- problem2_df %>%
  pull(vec_factor) %>%
  mean(na.rm = TRUE)

#vec_factor
mean_char <- problem2_df %>%
  pull(vec_char) %>%
  mean(na.rm = TRUE)
```
Answer: 
randomsample worked, it computed the average based on the numeric inputs. vec_char and vec_factor: You cannot calculate the mean of character or factor variables. If you attempt to calculate the mean of these types, you will get an error.


In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?
```{r}
print(as.numeric(problem2_df$vec_char))
print(as.numeric(problem2_df$vec_logical))
print(as.numeric(problem2_df$vec_factor))

```
What happened was when you convert a logical vector to numeric, the values TRUE/FALSE are converted to 1 and 0 respectively. When you convert a character vector to numeric, R cannot interpret the character strings as numbers, so it replaces them with NA values and spit out a warning: "NAs introduced by coercion." When you convert a factor vector to numeric, R returns the underlying integer codes that represent the factor levels, not the actual level names. In this case, high, low, medium are represented by 1, 2, 3. This happened because when you convert a factor to numeric, R gives you the internal integer codes representing the factor's levels, which may not have any meaningful interpretation in terms of the original factor labels. This does help explain what happens when I'm try to take the mean. Essentially, the mean of a factor converted to numeric is the average of the internal integer codes of the factor levels. This value can be misleading or meaningless because those numeric codes are simply R's way of organizing the factor levels—they don't reflect actual numeric values of the categories.
